Variance reduction for MCMC methods via martingale representations

### 2019-12-01 aap/Revision/vr-mcmc-revision.tex
Revision

\section{Concentration bounds for  ULA}
For $\gamma>0$, the Markov kernel $R_\gamma$
from $(\mathbb R^d,\mathcal B(\mathbb R^d))$
into $(\mathbb R^d,\mathcal B(\mathbb R^d))$
is defined by
$$
R_\gamma(x,A)=\int_{A} \frac{1}{(4 \pi \gamma)^{d/2}} \text{ exp} \left\{ -\frac{1}{4\gamma} \left\| y - x + \gamma \mu(x)\right\|^2\right\}dy,\quad A\in \mathcal B(\mathbb R^d),
$$
while, for $k,l\in\mathbb N$, $k\le l$,
the kernel $Q_{\boldsymbol{\gamma}}^{k,l}$ is
$
Q_{\boldsymbol{\gamma}}^{k,l}=R_{\gamma_l}\cdots R_{\gamma_k}.
$
 We make the following assumptions.

\begin{description}
\item[(H1)] Assume that $\nabla U$ is Lipschitz, that is, there exists $L\geq 0$ such that 
\begin{eqnarray*}
\|\nabla U(x)-\nabla U(y)\|\leq L\|x-y\|, \quad x,y\in \mathbb{R}^d
\end{eqnarray*}

@@

>>>>

\section{Bounds for moments of  ULA}
For $\gamma>0$, the Markov kernel $R_\gamma$
on $(\mathbb R^d,\mathcal B(\mathbb R^d))$
is defined by
$$
R_\gamma(x,A)=\int_{A} \frac{1}{(4 \pi \gamma)^{d/2}} \text{ exp} \left\{ -\frac{1}{4\gamma} \left\| y - x + \gamma \nabla U(x)\right\|^2\right\}dy,\quad A\in \mathcal B(\mathbb R^d),
$$
while, for $k,l\in\mathbb N$, $k\le l$,
the kernel $Q_{\boldsymbol{\gamma}}^{k,l}$ is
$
Q_{\boldsymbol{\gamma}}^{k,l}=R_{\gamma_l}\cdots R_{\gamma_k}.
$
 We make the following assumptions.

\begin{description}
\item[(H1)] Assume that $\nabla U$ is Lipschitz, that is, there exists $L\geq 0$ such that 
\begin{eqnarray*}
\|\nabla U(x)-\nabla U(y)\|\leq L\|x-y\|, \quad x,y\in \mathbb{R}^d.
\end{eqnarray*}

@@

\begin{thm}
Suppose that
\begin{eqnarray*}
\sum_{p=1}^{\infty} \gamma^2_p<\infty.
\end{eqnarray*}
Under assumptions (H1) and (H2), then for any bounded function $f$ on $\mathbb{R}^d$ 
we have
\begin{eqnarray*}
\mathsf{E}\left[\left(\sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X_{p})-\bar Q_l(X_l)\right)^{4}\right] \lesssim \frac{1}{\Gamma^4_{N+2,N+n+1}},
\end{eqnarray*}
 and
\begin{eqnarray*}
\bar Q_l(X_l)=\mathsf{E}\left[\left.\sum_{p=l}^{N+n}\omega_{p,n}^{N}f(X_{p})\right|X_{l}\right].
\end{eqnarray*}
Moreover 
\begin{eqnarray*}
\mathsf{E}\left[\exp\left(\rho\left(\sum_{k=l}^{p}\gamma_{k+1}g(X_{k})-\left(\sum_{k=l}^{p}\gamma_{k+1}\right)\pi(g)\right)\right)\right]\leq C
\end{eqnarray*}
for some small enough $\rho>0,$  all natural $0<l\leq p<\infty$ and constant $C$ not depending on $l,p.$
\end{thm}
